---
title: "R Notebook"
output: html_notebook
---

## Step 1 : 
#### Loading the Data
#### Plotting the data 
#### Conclusion

```{r}
tourism.data=readxl::read_xlsx("Tourism_new.xlsx")
head(tourism.data)
```

```{r}
dim(tourism.data)
```


```{r}
tail(tourism.data)
```


```{r}
# Monthly Seasonality (frequency set to 12 for monthly data)
tourism <- ts(Tourism_new[11:394,2],start= c(1988),end = c(2019,12),frequency=12)
tourism

```
```{r}
tourism=df2
```


```{r}
plot(tourism)
```
The graph clearly shows a seasonality and an increasing trend i.e. a multiplicative variance. So to remove the variance, to convert it into a additive series, we need to find the log returns of the series. Let us first plot the log returns of the series and see how it looks. 

```{r}
log_tourism <- log(tourism)
plot(log_tourism)
```
```{r}
Dlt_test_actual <- window(tourism, start= c(2018))
```


## Step 2 : Dummy fit => Ratio to MA => stationarity check!

#### a) Here we shall fit a dummy to find out if there is seasonality
#### b) If seasonality exists we will use Ratio to MA to remove the seasonality
#### c) then we shall check for non stochastic trend, and remove it finally.

##### a) Checking the existence of seasonality using dummy indices. 

```{r}
# all libraries required, some dependencies are there.
library(caret)
library(lattice)
library(ggplot2)
library(mltools)
library(data.table)

```
```{r results="hide" , message=FALSE, warningFALSE}
library(zoo)
as.yearmon(time(log_tourism)) #this statement finds the year and months

# Saving the months in  a vector
p <- month.abb[cycle(log_tourism)]

# Changing the months into factors
p <- as.factor(p)

# fitting dummy using mltools package
# the dummy variable is a list

dmy <- one_hot(as.data.table(p))
t <- 1:length(log_tourism)
# Setting the dummies and creating a dataframe
m <- as.matrix(log_tourism)
X <- cbind(as.data.frame(m), t, as.data.frame(dmy))
names(X)[1] <- "Y"
head(X, n= 10)

```
```{r}
summary(lm(Y~.,data = X))
```

The lm summary shows that many of the seasonal coefficients are significant showing that it has got a seasonality.
Next we apply *Ratio to MA filter*

```{r, message= FALSE , warning = FALSE}
library(astsa)
library(forecast)
MA <- ma(log_tourism, 12, centre = T)
id <- seq(1,length(log_tourism),1)
detrended <- log_tourism-MA
head(detrended, n = 10)
tail(detrended, n= 10)
```
```{r}
FTA_mat <- as.data.frame(matrix(detrended, nrow = 32, ncol= 12, byrow = T))
colnames(FTA_mat) <- month.name
rownames(FTA_mat) <- 1988:2019
S <- colMeans(FTA_mat, na.rm = T)
S_mean <- mean(S)
S_index <- S - S_mean
S_index
Deseasonalised_log_tourism <- log_tourism-S_index
Deseasonalised_log_tourism

```
```{r}
plot(log_tourism, main = "FTA from 1998 - 2019(log transformed)", ylab = "Number of tourists")
```
```{r}
plot(Deseasonalised_log_tourism, main = "Deseasonalised FTA 1998-2019 ", ylab = "Number FT")

```
```{r}
plot(log_tourism, main = "FTA from 1988 - 2019(original = black, Deseasoned = Red)", ylab = "Number of tourists")
lines(Deseasonalised_log_tourism, col = "red")
```
```{r}
FTA_DL_train <-  window(Deseasonalised_log_tourism, end = c(2017,12))
FTA_DL_test <- window(Deseasonalised_log_tourism, start= c(2018))
plot(FTA_DL_train)
lines(FTA_DL_test, col = "red")
```
```{r}
library(urca)
summary(ur.df(Deseasonalised_log_tourism, type = "trend", lags = 15, selectlags = "AIC"))
```
```{r}
diff1_Deaseasoned_log_tourism <- diff(Deseasonalised_log_tourism)
```
```{r}
library(urca)
summary(ur.df(diff1_Deaseasoned_log_tourism, type = "trend", lags = 15, selectlags = "AIC"))
```
This shows that the modulus of the test statistic is 2.08 < 3.98 so we can say that our data has got a unit root, i.e.stochastic trend which needs to be removed first.

For that we take the first lag difference.

```{r}
diff1_Deaseasoned_log_tourism <- diff(Deseasonalised_log_tourism)
```

Now again we shall check if there is still a stochastic trend. 
We will confirm the same using the *Augmented Dickey Fuller test*

```{r}
library(urca)
summary(ur.df(diff1_Deaseasoned_log_tourism, type = "trend", lags = 15, selectlags = "AIC"))
```

This shows that the modulus of the test statistic is 9.6332 > 3.98 so we can say that our data has no unit root, i.e. stochastic trend is not there any more.

*Note:* Here the function has automatically chosen the lag which takes the lowest AIC, and calculates accordingly.

Now we fit a linear trend to the data to check if there is any deterministic trend.

```{r}
model_select <- function(y, x ,poly.deg)
{
  model <- lm(y~poly(x,poly.deg))
  return(summary(model))
  
}
t=seq.int(1,length(diff1_Deaseasoned_log_tourism),1)
y <- diff1_Deaseasoned_log_tourism
```

Fitting a trend model: $y_t = a+ bt + u_t$

```{r}
model_select(y,t,1)
```

Well this shows that we can fail to reject the null hypothesis that is the coefficient b = 0, so here apparently there does not exist a deterministic trend as such in the data. 

Now that the process is stationary from a *Trend Stationary process*, we proceed forward to building the model. 

## Step 4 : Sample Dividing Procedure. 

Now we shall divide the data into *In sample* and *Out Sample*, we shall keep 30 years for the training (in) sample and rest 2 years for forecasting (out sample). 

```{r}
Dlt_train <-  window(diff1_Deaseasoned_log_tourism, end = c(2017,12))
Dlt_test <- window(diff1_Deaseasoned_log_tourism, start= c(2018))
plot(Dlt_train)
lines(Dlt_test,col= "red")
```
```{r}
set.seed(1)
library(nnfor)
fit_mlp=mlp(FTA_DL_train,lags=1:12,hd.auto.type="valid",hd.max=15,reps = 50, outplot = T)
print(fit_mlp)
```


```{r}
print(fit_mlp)
```
```{r}
plot(fit_mlp)
```

```{r}
for_mlp= forecast(fit_mlp,h=24)
print(for_mlp)
```
```{r}
PRED_ACTtable <- data.frame(Predict = as.numeric(for_mlp$mean), as.numeric(FTA_DL_test))
colnames(PRED_ACTtable) <- c("Predict", "Actual")
PRED_ACTtable
```

```{r}
rmse_DL <- sqrt(mean((PRED_ACTtable$Predict- PRED_ACTtable$Actual)^2))
rmse_DL
mape_DL <- mean(abs(PRED_ACTtable$Predict- PRED_ACTtable$Actual)/abs(PRED_ACTtable$Actual))*100
mape_DL
```
```{r}
plot(PRED_ACTtable$Actual)
lines(PRED_ACTtable$Predict, col = "red")
```


```{r}
Predict <- exp(PRED_ACTtable$Predict+S_index)
Predict
Actual <- window(tourism, start= c(2018), end = c(2019,12))
Actual_predict_table <- data.frame(Predict = Predict, Actual = Actual)
Actual_predict_table
```


```{r}
plot(Actual)
lines(Predict, col = "red")
```

```{r}
rmse <- sqrt(mean((Predict-Actual)^2))
rmse
mape <- mean(abs(Predict- Actual)/abs(Actual))*100
mape
```
```{r}
plot(fit_mlp$MSEH, ylab="Mean Squared Error", xlab="Number of neurons")
lines(fit_mlp$MSEH)
```
```{r}
fit_mlp$MSEH
```

